Reproducible Research Project 1


Output: 
  html_document:
    keep_md: true
    
    
## Loading and preprocessing the data

```{r}

setwd("./R/Reproducible Research")

activityData <- read.csv(file = "activity.csv", sep =",", colClasses=c("integer","Date","integer"))
str(activityData)
summary(activityData)

```

## What is Mean Total Number of Steps Taken Per Day

Mean Number of Steps

```{r}

totalSteps <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
stepMean <- mean(totalSteps)
stepMean

Median 

stepMedian <- median(totalSteps)
stepMedian

```
## Histogram of Total Steps Taken Each Day

```{r}

hist(totalSteps, breaks = 12,
     xlab = "Number of Steps Per Day",
     ylab = "Frequency of Occurance in Data",
     main = "Histogram of total steps per day")

abline(v=stepMean, col="purple", lwd ="4")
abline(v=stepMedian, col="brown", lwd ="4")
legend(x = "topright", legend = c("mean", "median"), col=c("purple","brown"), lwd = 4)

```

## Average Daily Activity Pattern

```{r}

library(plyr)
dailyActivity <- ddply(activityData, .(interval), summarize, steps = mean(steps, na.rm = TRUE))
with(dailyActivity, plot(interval, steps, type = "l"))

```

5 Minutes Containing the Maximum Number of Steps

```{r}

maxSteps <- dailyActivity[which.max(dailyActivity$steps),]$interval

```

## Inputing Missing Values

Find Out How Many Missing Values


```{r}

sum(is.na(activityData$steps))

```

# Filling in All the Missing Values in the Dataset

```{r}


stepValues <- data.frame(activityData$steps)
stepValues [is.na(stepValues),] <- ceiling(tapply(X=activityData$steps,INDEX=activityData$interval,FUN=mean,na.rm=TRUE))

newData <- cbind(stepValues, activityData[,2:3])
colnames(newData) <- c("Steps", "Date", "Interval")

summary(newData)


```

Newly Inputed Dataset from the Old Dataset

```{r}

totalSteps2 <- tapply(newData$Steps, newData$Date, sum, na.rm=TRUE)
stepMean2 <- mean(totalSteps2)
stepMean2

stepMedian2 <- median(totalSteps2)
stepMedian2

```


## New Input Histogram Data

```{r}

hist(totalSteps2, breaks = 12,
     xlab = "Number of Steps Per Day",
     ylab = "Frequency of Occurance in Data",
     main = "Histogram of total steps per day")

abline(v=stepMean2, col="purple", lwd ="5")
abline(v=stepMedian2, col="brown", lwd ="3")
legend(x = "topright", legend = c("mean", "median"), col=c("purple","brown"), lwd = 4)

```

## Differences in Activity Patterns Between Weekend and Weekdays

```{r}

newData$Weekend <- weekdays(newData$Date) == "Saturday" | weekdays(newData$Date) == "Sunday"

newData$Weekend <- factor(newData$Weekend, levels = c(F, T), labels = c("Weekday", "Weekend"))

activity <- ddply(newData, .(Interval, Weekend), summarize, steps = mean(Steps, na.rm = TRUE))

library(lattice)
xyplot(steps ~ Interval | Weekend, activity, type = "l", layout = c(1, 2), ylab = "Number of Steps", xlab = "Interval", main = "Weekend vs. Weekday activity patterns")

```

